Non-word repetition in Yélî Dnye: A small-scale study

### 2020-05-14 manuscript.Rmd
lots of bug fixes

```
## TODO

- create stimuli table that has complexity & length
- integrate demo data & stim table in final data
- add analysis on the basis of complexity & length
- add analysis on the basis of demo chars
- fix the back-transformation
- add summary of types of error: %semantic, %deletion (which sounds?) %subst (which are the most common sounds) 
- maybe add to the last ana the prevalence (ie out of all the gh which proportion get transformed)
- discussion

@@

```{r readin}
read.table("final_data.txt",header=T,encoding="UTF-8")->trnsc
#create a subset with just the first presentations -- this is what is going to be used for the quantitative analyses
firstp=subset(trnsc,rep==1 & !is.na(rep))
```

Taking into account only the first attempts, the overall repetition accuracy measured in whole words is `r round(mean(firstp$cor.online)*100)`% based on the online coding (i.e., whether the native experimenter asked for the target to be re-played), and `r round(mean(firstp$correct)*100)`% when the offline coding is considered. On average, the phoneme-based normalized Levenshtein distance was `r round(mean(firstp$nld)*100)`%, meaning that about 20% of phonemes were substituted, inserted, or deleted.

```{r figlength,eval=F}
#not working
aggregate(firstp$correct,list(firstp$tarlen,firstp$id),mean)
plot(firstp$correct~firstp$tarlen)

```
- How frequent are errors that result in real words? Is that a function of item complexity? What are some common patterns of phonological errors?

@@

@@

@@

We start by describing whole word accuracy. Participants scored

@@

phoneme based, distance

### Accuracy a function of item complexity 
(number of syllablex, sound complexity)?

>>>>

## TODO

- fix yaml
- add ana based on phonemes & expand on levenshtein
- add ana based on number of characters
- add summary of types of error: %semantic, %deletion (which sounds?) %subst (which are the most common sounds) 
- maybe add to the last ana the prevalence (ie out of all the gh which proportion get transformed)
- add analysis on the basis of demo chars
- discussion

@@

@@

```{r readin}
read.table("final_data.txt",header=T,encoding="UTF-8")->trnsc

@@

#remove children who do not have both first & subsequent attempts
pwcm_att[pwcm_att$ID %in% names(table(pwcm_att$ID))[table(pwcm_att$ID)==2],]->pwcm_att

@@

Taking into account only the first attempts, the overall repetition accuracy measured in whole words is `r round(mean(trnsc$cor.online)*100)`% based on the online coding (i.e., whether the native experimenter asked for the target to be re-played), and `r round(mean(trnsc$correct)*100)`% when the offline coding is considered. On average, the phoneme-based normalized Levenshtein distance was `r round(mean(trnsc$nld)*100)`%, meaning that about 20% of phonemes were substituted, inserted, or deleted.

ADD GRAPH WITH PHONEME BASED IN Y AND WORD BASED IN X

### Accuracy a function of item complexity

```{r figlength,eval=F}
#not working
aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$age.exact,trnsc$tarlen),mean)->pwcm_att
colnames(pwcm_att)<-c("ID","Age","Length","Word")

#but this one is! What a mystery...
aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$age.exact,trnsc$type),mean)->pwcm_att
colnames(pwcm_att)<-c("ID","Age","Length","Word")

myx=as.numeric(as.factor(pwcm_att$ID))/100+as.numeric(as.factor(pwcm_att$Length))

plot(pwcm_att$Word~myx,pch=20,cex=2,lty=3,xlab="",xaxt="n",ylab="NWR scores (%)")
axis(1,at=1:length(levels(factor(pwcm_att$Length))),labels=levels(factor(pwcm_att$Length)))

@@

### 2020-05-14 manuscript.tex
lots of bug fixes

@@

The child was donned a headset (\textbf{xx ({\textbf{???}}) please fill in} for most of the children, xx ({\textbf{???}}) fill in for the rest). For most children, the headset could not stay comfortably on the child's head, and thus it was placed on the child's shoulders, with the microphone carefully placed close to the child's mouth. A local informant sat next to the child, to provide the instructions and, if needed, coach the child to make sure that they understood that the task was to repeat the items precisely without changing anything using the three practice items as well as real words. The experimenter (the first author) delivered the elicitation stimuli to the local informant and the child over headhphones.

@@

A total 55 children were tested, from 34 of families, in five hamlets. Some children could not be included for the following reasons: refused participation or failed to repeat items presented over headphones even after coaching (N=8), spoke too softly to allow offline coding (N=5). In addition, 2 teenagers were tested to put younger children at ease; their data is not included in analyses below. The remaining 40 children (14 girls) were aged 6.96 years (range 3.92-11.03 years). There were 32 children exposed only to Yélî in the home, 6 children who were also exposed to another language in the home, and , 2 for whom this information was missing. Maternal years of education averaged 8.24 years (range 6-12 years; 2 children had this information missing).\footnote{Education is often reported in even years because people typically complete two-year cycles.} In terms of birth order, 0 were first borns, 5 second, 4 third, 2 forth, 6 fifth, 5 sixth, and 1 did not have this information.

@@

@@

We start by describing whole word accuracy

phoneme based, distance
\#\#\# Accuracy a function of item complexity
(number of syllablex, sound complexity)?

>>>>

\hypertarget{todo}{%
\subsection{TODO}\label{todo}}

\begin{itemize}
\tightlist
\item
  fix yaml
\item
  add ana based on phonemes \& expand on levenshtein
\item
  add ana based on number of characters
\item
  add summary of types of error: \%semantic, \%deletion (which sounds?) \%subst (which are the most common sounds)
\item
  maybe add to the last ana the prevalence (ie out of all the gh which proportion get transformed)
\item
  add analysis on the basis of demo chars
\item
  discussion
\end{itemize}

@@

The child was donned a headset (\textbf{xx ({\textbf{???}}) please fill in} for most of the children, SHURE WH20 XLR headset with a dynamic microphone for the rest), recorded into the left channel into a Tascam DR40x digital recorder. For most children, the headset could not stay comfortably on the child's head, and thus it was placed on the child's shoulders, with the microphone carefully placed close to the child's mouth. A local informant sat next to the child, to would provide the instructions and, if needed, coach the child to make sure, using the three practice items as well as real words, that they understood that the task was to repeat the items precisely without changing anything. An experimenter (the first author) delivered the elicitation stimuli to the local informant and the child over headphones.

@@

A total 55 children were tested, from 34 of families, in five hamlets. Some children could not be included for the following reasons: refused participation or failed to repeat items presented over headphones even after coaching (N=0), spoke too softly to allow offline coding (N=5). In addition, 2 teenagers were tested to put younger children at ease; their data is not included in analyses below. The remaining 40 children (14 girls) were aged 6.96 years (range 3.92-11.03 years). There were 32 children exposed only to Yélî in the home, 6 children who were also exposed to another language in the home, and , 2 for whom this information was missing. Maternal years of education averaged 8.24 years (range 6-12 years; 2 children had this information missing).\footnote{Education is often reported in even years because people typically complete two-year cycles.} In terms of birth order, 0 were first borns, 5 second, 4 third, 2 forth, 6 fifth, 5 sixth, and 1 did not have this information.

@@

\hypertarget{preliminary-analyses}{%
\subsubsection{Preliminary analyses}\label{preliminary-analyses}}

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/Fig1-first_vs_others-1.pdf}
\caption{(\#fig:Fig1-first\_vs\_others)NWR scores for individual participants averaging separately their first attempts and all other attempts.}
\end{figure}

We first checked whether accuracy varies between first and subsequent presentations of an item by averaging word-level accuracy at the participant level separately for first attempts and subsequent repetitions. As shown in Figure 1, participants' mean word-level accuracy became more heterogeneous in subsequent repetitions. Surprisingly, subsequent repetitions (M = 42, SD = 28)
were on average less accurate than first ones (M = 64, SD = 15), t(36) = 5.32, p = 0. Given the uncertainty in whether previous work used only the first or all repetitions, and since behavior degraded and became more heterogeneous in subsequent repetitions, the rest of the analyses focus on only the first repetitions.

@@

Taking into account only the first attempts, the overall repetition accuracy measured in whole words is 72\% based on the online coding (i.e., whether the native experimenter asked for the target to be re-played), and 65\% when the offline coding is considered. On average, the phoneme-based normalized Levenshtein distance was 22\%, meaning that about 20\% of phonemes were substituted, inserted, or deleted.

ADD GRAPH WITH PHONEME BASED IN Y AND WORD BASED IN X

\hypertarget{accuracy-a-function-of-item-complexity}{%
\subsubsection{Accuracy a function of item complexity}\label{accuracy-a-function-of-item-complexity}}

We then inspected whether accuracy varied as a function of word length.
